---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
In order to load the data we need to use the *read.csv* function.

```{r}
wd <- getwd()
raw_data <- read.csv(file=file.path(wd,"data", "activity.csv"))
```

As the raw data does not provide the total number of steps per day, rather it presents the number of steps on a 5-second-interval basis, it is necessary to calculate the aggregate sum of steps.

```{r}
daily_steps <- aggregate(
  steps ~ date,
  data = raw_data,
  FUN = sum,
  na.rm = TRUE
)
```

## What is mean total number of steps taken per day?
### 1.Make a histogram of the total number of steps taken each day

```{r}
hist(
  daily_steps$steps,
  main = "Histogram of Steps",
  xlab = "Steps",
  ylab = "Frequency"
)
```

```{r}
steps_mean <- mean(daily_steps$steps)
steps_median <- median(daily_steps$steps)
```
### 2. Calculate and report the mean and median total number of steps taken per day

The mean total number of steps per days is `r steps_mean` steps.  
The median total number of steps per days is `r steps_median` steps.

## What is the average daily activity pattern?

## Imputing missing values

## Are there differences in activity patterns between weekdays and weekends?
