---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
In order to load the data we need to use the *read.csv* function.

```{r}
wd <- getwd()
raw_data <- read.csv(file=file.path(wd,"data", "activity.csv"))
```

As the raw data does not provide the total number of steps per day, rather it presents the number of steps on a 5-minute interval basis, it is necessary to calculate the aggregate sum of steps.

```{r}
daily_steps <- aggregate(
  steps ~ date,
  data = raw_data,
  FUN = sum,
  na.rm = TRUE
)
```

## What is mean total number of steps taken per day?
### 1.Make a histogram of the total number of steps taken each day

```{r}
hist(
  daily_steps$steps,
  main = "Histogram of Steps",
  xlab = "Steps",
  ylab = "Frequency"
)
```

### 2. Calculate and report the mean and median total number of steps taken per day

```{r}
steps_mean <- mean(daily_steps$steps)
steps_median <- median(daily_steps$steps)
```
The mean total number of steps per days is `r steps_mean` steps.  
The median total number of steps per days is `r steps_median` steps.

## What is the average daily activity pattern?
### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

We need to again use the raw data and aggregate the number of steps per each 5-minute interval across all days:  
```{r}
avg_steps <- aggregate(
  steps ~ interval,
  data = raw_data,
  FUN = mean,
  na.rm = TRUE
)
```
After preparing a new data frame, we can proceed with plotting the data:

```{r}
plot(
  avg_steps$interval,
  avg_steps$steps,
  type = "l",
  xlab = "Interval",
  ylab = "Average Number of Steps",
  main = "Average Daily Activity Pattern"
)
```
### 2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_mean_steps <- max(avg_steps$steps)
max_mean_steps_interval <- avg_steps$interval[
  avg_steps$steps == max_mean_steps
]
```
The interval with the highest average number of steps is number `r max_mean_steps_interval` with `r max_mean_steps` steps.  

## Imputing missing values

## Are there differences in activity patterns between weekdays and weekends?
